ANCOVA Comparison Simulations: Nonlinear Model

### 2017-10-09 simulations/continuous_outcomes.Rmd
Create power ratio plots for the simulations

@@

```{r vary_effect_size_plot, fig.height=4}
load("continuous_outcomes/vary_gamma_results.Rda")
gamma_vec <- seq(0, 0.2, by=0.05)
gamma_power <- lapply(gamma_res, function(x) apply(x, 2, compute_power))
plot_power_gamma(gamma_power, tr_effect, 0.05, "Power at level 5%, Constant Treatment Effect") 
```

@@

# Gaussian error terms, heterogeneous treatment effect
```{r vary_effect_size_het_plot, fig.height=4}
load("continuous_outcomes/vary_gamma_het_results.Rda")
gamma_vec <- seq(0, 0.2, by=0.05)
gamma_power_het <- lapply(gamma_res_het, function(x) apply(x, 2, compute_power))
plot_power_gamma(gamma_power_het, tr_effect_het, 0.05, "Power at level 5%, Heterogeneous Treatment Effect") 
```
Next, we varied the distribution of $v_{ij}$ across strata and generated normally distributed errors.
Once again, we varied $\gamma$ from $0$ to $0.2$ in steps of $0.05$.
The corresponding population average treatment effects were of `r round(tr_effect_het, 2)[1:4]`, and `r round(tr_effect_het, 2)[5]`, respectively.
Figure 2 shows the empirical power at level 5\% for these increasing effect sizes.
Four of the five tests had the correct level of 5\%, while the stratified permutation test rejected only 1\% of the time.
Both the stratified and differenced permutation tests had lower power than the linear model tests as the effect size increased.
The correlation between baseline $X$ and outcome $Y$ was low but nontrivial (between 0.04 and 0.44), so this result makes some sense:
it is beneficial for power to control for the baseline in the linear model, but the correlation is too low to make the differencing method useful.

@@

```{r plot_for_ms, include=FALSE}
p1 <- plot_power_gamma(gamma_power, tr_effect, 0.05, "Power at level 5%, Constant Treatment Effect") + 
  ylim(0, 0.9) +
  theme(legend.position="bottom") + 
  guides(color=guide_legend(nrow=2,byrow=TRUE)) +
  report_theme +
  ggtitle("Constant Treatment \n Effect")

p2 <- plot_power_gamma(gamma_power_het, tr_effect_het, 0.05, "Power at level 5%, Heterogeneous Treatment Effect") +
    ylim(0, 0.9) +
    report_theme +
    ggtitle("Heterogeneous Treatment \n Effect")

>>>>

plot_power_ratio_gamma <- function(powermat_list, gamma_vec, alpha, title){
  gamma_power_alpha <- t(sapply(powermat_list, function(x) x[floor(alpha*nrow(x)),]))
  colnames(gamma_power_alpha) <- c("ANCOVA", "Stratified Permutation", "Differenced Permutation", "LM Permutation", "Freedman-Lane")
  gamma_power_alpha <- apply(gamma_power_alpha, 2, function(x) x/gamma_power_alpha[,"ANCOVA"])
  melt(gamma_power_alpha, value.name="Power") %>%
  mutate("Method" = Var2) %>%
  mutate("Effect" = rep(gamma_vec, 5)) %>% 
  filter(Method != "Differenced Permutation") %>%
  filter(Method != "ANCOVA") %>%
  ggplot(aes(x = Effect, y = Power)) +
  geom_line(aes(color = Method))+
  xlab("Effect size") +
  ylab("Relative Power") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 16),
    title = element_text(size = 16),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 14),
    strip.text.x = element_text(size = 12)
  ) 
}
```

@@

```{r vary_effect_size_plot, fig.height=4}
load("continuous_outcomes/vary_gamma_results.Rda")
gamma_vec <- seq(0, 0.2, by=0.05)
gamma_power <- lapply(gamma_res, function(x) apply(x, 2, compute_power))
plot_power_gamma(gamma_power, tr_effect, 0.05, "Power at level 5%, Constant Treatment Effect") 
plot_power_ratio_gamma(gamma_power, tr_effect, 0.05, "Power at level 5%, Constant Treatment Effect")
```

@@

# Gaussian error terms, heterogeneous treatment effect
```{r vary_effect_size_het_plot, fig.height=4}
load("continuous_outcomes/vary_gamma_het_results.Rda")
gamma_vec <- seq(0, 0.2, by=0.05)
gamma_power_het <- lapply(gamma_res_het, function(x) apply(x, 2, compute_power))
plot_power_gamma(gamma_power_het, tr_effect_het, 0.05, "Power at level 5%, Heterogeneous Treatment Effect")

plot_power_ratio_gamma(gamma_power_het, tr_effect_het, 0.05, "Power at level 5%, Heterogeneous Treatment Effect")

```
Next, we varied the distribution of $v_{ij}$ across strata and generated normally distributed errors.
Once again, we varied $\gamma$ from $0$ to $0.2$ in steps of $0.05$.
The corresponding population average treatment effects were of `r round(tr_effect_het, 2)[1:4]`, and `r round(tr_effect_het, 2)[5]`, respectively.
Figure 2 shows the empirical power at level 5\% for these increasing effect sizes.
Four of the five tests had the correct level of 5\%, while the stratified permutation test rejected only 1\% of the time.
Both the stratified and differenced permutation tests had lower power than the linear model tests as the effect size increased.
The correlation between baseline $X$ and outcome $Y$ was low but nontrivial (between 0.04 and 0.44), so this result makes some sense:
it is beneficial for power to control for the baseline in the linear model, but the correlation is too low to make the differencing method useful.

@@

```{r plot_for_ms, include=FALSE}
p1 <- plot_power_ratio_gamma(gamma_power, tr_effect, 0.05, "Power at level 5%, Constant Treatment Effect") + 
  ylim(0, 1.2) +
  theme(legend.position="bottom") + 
  guides(color=guide_legend(nrow=1,byrow=TRUE)) +
  report_theme +
  ggtitle("Constant Treatment \n Effect")

p2 <- plot_power_ratio_gamma(gamma_power_het, tr_effect_het, 0.05, "Power at level 5%, Heterogeneous Treatment Effect") +
  ylim(0, 1.2) +
    report_theme +
    ggtitle("Heterogeneous Treatment \n Effect")

### 2017-10-09 simulations/nonlinear_model.Rmd
Create power ratio plots for the simulations

@@

```{r vary_effect_size_plot, fig.height=4}
load("nonlinear_model/vary_gamma_results.Rda")
gamma_vec <- seq(0, 0.2, by=0.05)
gamma_power <- lapply(gamma_res, function(x) apply(x, 2, compute_power))
plot_power_gamma(gamma_power, tr_effect, 0.05, "Power at level 5%, Constant Treatment Effect") 
```

@@

# Heterogeneous treatment effect
```{r vary_effect_size_het_plot, fig.height=4}
load("nonlinear_model/vary_gamma_het_results.Rda")
gamma_vec <- seq(0, 0.2, by=0.05)
gamma_power_het <- lapply(gamma_res_het, function(x) apply(x, 2, compute_power))
plot_power_gamma(gamma_power_het, tr_effect_het, 0.05, "Power at level 5%, Heterogeneous Treatment Effect") 
```
Next, we varied the distribution of $v_{ij}$ across strata and generated normally distributed errors.
Once again, we varied $\gamma$ from $0$ to $0.2$ in steps of $0.05$.
The corresponding population average treatment effects were of `r round(tr_effect_het, 2)[1:4]`, and `r round(tr_effect_het, 2)[5]`, respectively.
Figure 2 shows the empirical power at level 5\% for these increasing effect sizes.
Four of the five tests had the correct level of 5\%, while the stratified permutation test rejected only 1\% of the time.
The same pattern appeared here: the stratified permutation test rarely rejects, while the differenced permutation test dominates all of the linear model based tests.

>>>>

plot_power_ratio_gamma <- function(powermat_list, gamma_vec, alpha, title){
  gamma_power_alpha <- t(sapply(powermat_list, function(x) x[floor(alpha*nrow(x)),]))
  colnames(gamma_power_alpha) <- c("ANCOVA", "Stratified Permutation", "Differenced Permutation", "LM Permutation", "Freedman-Lane")
  gamma_power_alpha <- apply(gamma_power_alpha, 2, function(x) x/gamma_power_alpha[,"ANCOVA"])
  melt(gamma_power_alpha, value.name="Power") %>%
  mutate("Method" = Var2) %>%
  mutate("Effect" = rep(gamma_vec, 5)) %>% 
  filter(Method != "Differenced Permutation") %>%
  filter(Method != "ANCOVA") %>%
  ggplot(aes(x = Effect, y = Power)) +
  geom_line(aes(color = Method))+
  xlab("Effect size") +
  ylab("Relative Power") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 16),
    title = element_text(size = 16),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 14),
    strip.text.x = element_text(size = 12)
  ) 
}

@@

```{r vary_effect_size_plot, fig.height=4}
load("nonlinear_model/vary_gamma_results.Rda")
gamma_vec <- seq(0, 0.2, by=0.05)
gamma_power <- lapply(gamma_res, function(x) apply(x, 2, compute_power))
plot_power_gamma(gamma_power, tr_effect, 0.05, "Power at level 5%, Constant Treatment Effect") 
plot_power_ratio_gamma(gamma_power, tr_effect, 0.05, "Power at level 5%, Constant Treatment Effect")

@@

# Heterogeneous treatment effect
```{r vary_effect_size_het_plot, fig.height=4}
load("nonlinear_model/vary_gamma_het_results.Rda")
gamma_vec <- seq(0, 0.2, by=0.05)
gamma_power_het <- lapply(gamma_res_het, function(x) apply(x, 2, compute_power))
plot_power_gamma(gamma_power_het, tr_effect_het, 0.05, "Power at level 5%, Heterogeneous Treatment Effect")

plot_power_ratio_gamma(gamma_power_het, tr_effect_het, 0.05, "Power at level 5%, Heterogeneous Treatment Effect") 
```
Next, we varied the distribution of $v_{ij}$ across strata and generated normally distributed errors.
Once again, we varied $\gamma$ from $0$ to $0.2$ in steps of $0.05$.
The corresponding population average treatment effects were of `r round(tr_effect_het, 2)[1:4]`, and `r round(tr_effect_het, 2)[5]`, respectively.
Figure 2 shows the empirical power at level 5\% for these increasing effect sizes.
Four of the five tests had the correct level of 5\%, while the stratified permutation test rejected only 1\% of the time.
The same pattern appeared here: the stratified permutation test rarely rejects, while the differenced permutation test dominates all of the linear model based tests.
